---
title: "R Notebook"
output: html_notebook
---

#Chapter 5 The tidyverse

Once we start more advanced analyses, the preferred unit for data storage is not the vector but the data frame. In this chapter we learn to work directly with data frames, which greatly facilitate the organization of information.. We will focus on a specific data format referred to as tidy and on specific collection of packages that are particularly helpful for working with tidy data referred to as the tidyverse.

We can load all the tidyverse packages at once by installing and loading the **tidyverse** package:

```{r}
library(tidyverse)
```

We will learn how to implement the tidyverse approach throughout the book, but before delving into the details, in this chapter we introduce some of the most widely used tidyverse functionality. starting with the **dplyr** package for manipulating data frames and the **purrr** package for working with functions. Note that the tidyverse also includes a graphing package, **ggplot2**, which we introduce later in Chapter 8 in the Data Visualization part of the book; the **readr** package discussed in Chapter 6; and many others. In this chapter, we first introduce the concept of tidy data and then demonstrate how we use the tidyverse to work with data frames in this format.




## 5.1 Tidy data

We say that a data table is in tidy format if each row represents one observation and columns represent the different variables available for each of these observations. The `murders` dataset is an example of a tidy data frame.

```{r}
#>        state abb region population total
#> 1    Alabama  AL  South    4779736   135
#> 2     Alaska  AK   West     710231    19
#> 3    Arizona  AZ   West    6392017   232
#> 4   Arkansas  AR  South    2915918    93
#> 5 California  CA   West   37253956  1257
#> 6   Colorado  CO   West    5029196    65
```

Each row represent a state with each of the five columns providing a different variable related to these states: name, abbreviation, region, population, and total murders.

To see how the same information can be provided in different formats, consider the following example:

```{r}
#>        country year fertility
#> 1      Germany 1960      2.41
#> 2  South Korea 1960      6.16
#> 3      Germany 1961      2.44
#> 4  South Korea 1961      5.99
#> 5      Germany 1962      2.47
#> 6  South Korea 1962      5.79
#> 7      Germany 1963      2.49
#> 8  South Korea 1963      5.57
#> 9      Germany 1964      2.49
#> 10 South Korea 1964      5.36
#> 11     Germany 1965      2.48
#> 12 South Korea 1965      5.16
```

This tidy dataset provides fertility rates for two countries across the years. This is a tidy dataset because each row presents one observation with the three variables being county, year and fertility rate. However, this dataset originally came in another format and was reshaped for the **dslabs** package. Originally, the data was in the following format:

```{r}
#>       country 1960 1961 1962 1963 1964 1965
#> 1     Germany 2.41 2.44 2.47 2.49 2.49 2.48
#> 2 South Korea 6.16 5.99 5.79 5.57 5.36 5.16
```

The same information is provided, but there are two important differences in the format: 1) each row includes several observations and 2) one of the variables, year, is stored in the header. For the tidyverse packages to be optimally used, data need to be reshaped into `tidy` format, which you will learn to do in the Data Wrangling part of the book. Until then, we will use example datasets that are already in tidy format.

Although not immediately obvious, as you go through the book you will start to appreciate the advantages of working in a framework in which functions use tidy formats for both inputs and outputs. You will see how this permits the data analyst to focus on more important aspects of the analysis rather than the format of the data.



## 5.2 Exercises

1. Examine the built-in dataset `co2`. Which of the following is true:

```{r}
co2
```


A. `co2` is tidy data: it has one year for each row.

B. `co2` is not tidy: we need at least one column with a character vector.

C. `co2` is not tidy: it is a matrix instead of a data frame.

**D**. `co2` is not tidy: to be tidy we would have to wrangle it to have three columns (year, month and value), then each co2 observation would have a row.

??



2. Examine the built-in dataset `ChickWeight`. Which of the following is true:

```{r}
ChickWeight
```


A. `ChickWeight` is not tidy: each chick has more than one row.

**B**. `ChickWeight` is tidy: each observation (a weight) is represented by one row. The chick from which this measurement came from is one the variables.

C. `ChickWeight` is not a tidy: we are missing the year column.

D. `ChickWeight` is tidy: it is stored in a data frame.

??



3. Examine the built-in dataset `BOD`. Which of the following is true:

```{r}
BOD
```


A. `BOD` is not tidy: it only has six rows.

B. `BOD` is not tidy: the first column is just an index.

**C**. `BOD` is tidy: each row is an observation with two values (time and demand)

D. `BOD` is tidy: all small datasets are tidy by definition.

??



4. Which of the following built-in datasets is tidy (you can pick more than one):

```{r}
UCBAdmissions
```


A. `BJsales`

B. `EuStockMarkets`

**C**. `DNase`

**D**. `Formaldehyde`

**E**. `Orange`

F. `UCBAdmissions`



## 5.3 Manipulating data frames

The *dplyr* package from the *tidyverse* introduces functions that perform some of the most common operations when working with data frames and uses names for these functions that are relatively easy to remember. For instance, to change the data table by adding a new column, we use `mutate`. To filter the data table to a subset of rows, we use `filter`. Finally, to subset the data by selecting specific columns, we use `select`.



### 5.3.1 Adding a column with `mutate`

The function `mutate` takes the data frame as a first argument and the name and values of the variable as a second argument using the convention `name = values`. So, to add murder rates, we use:

```{r}
library(dslabs)
data("murders")
murders <- mutate(murders, rate = total / population * 100000)
```

Notice that here we used `total` and `population` inside the function, which are objects that are not defined in our workspace. But why don’t we get an error?

This is one of *dplyr*’s main features. Functions in this package, such as `mutate`, know to look for variables in the data frame provided in the first argument. In the call to mutate above, `total` will have the values in `murders$total`.

We can see that the new column is added:

```{r}
head(murders)
#>        state abb region population total rate
#> 1    Alabama  AL  South    4779736   135 2.82
#> 2     Alaska  AK   West     710231    19 2.68
#> 3    Arizona  AZ   West    6392017   232 3.63
#> 4   Arkansas  AR  South    2915918    93 3.19
#> 5 California  CA   West   37253956  1257 3.37
#> 6   Colorado  CO   West    5029196    65 1.29
```

Although we have overwritten the original `murders` object, this does not change the object that loaded with `data(murders)`. If we load the `murders` data again, the original will overwrite our mutated version.



### 5.3.2 Subsetting with `filter`

Now suppose that we want to filter the data table to only show the entries for which the murder rate is lower than 0.71. To do this we use the `filter` function, which takes the data table as the first argument and then the conditional statement as the second. Like `mutate`, we can use the unquoted variable names from `murders` inside the function and it will know we mean the columns and not objects in the workspace.

```{r}
filter(murders, rate <= 0.71)
#>           state abb        region population total  rate
#> 1        Hawaii  HI          West    1360301     7 0.515
#> 2          Iowa  IA North Central    3046355    21 0.689
#> 3 New Hampshire  NH     Northeast    1316470     5 0.380
#> 4  North Dakota  ND North Central     672591     4 0.595
#> 5       Vermont  VT     Northeast     625741     2 0.320
```



### 5.3.3 Selecting columns with `select`

Although our data table only has six columns, some data tables include hundreds. If we want to view just a few, we can use the *dplyr* `select` function. In the code below we select three columns, assign this to a new object and then filter the new object:

```{r}
new_table <- select(murders, state, region, rate)
filter(new_table, rate <= 0.71)
#>           state        region  rate
#> 1        Hawaii          West 0.515
#> 2          Iowa North Central 0.689
#> 3 New Hampshire     Northeast 0.380
#> 4  North Dakota North Central 0.595
#> 5       Vermont     Northeast 0.320
```

In the call to `select`, the first argument `murders` is an object, but `state`, `region`, and `rate` are variable names.



## 5.4 Exercises

1. Load the **dplyr** package and the murders dataset.

```{r}
library(dplyr)
library(dslabs)
data(murders)
```

You can add columns using the **dplyr** function `mutate`. This function is aware of the column names and inside the function you can call them unquoted:

```{r}
murders <- mutate(murders, population_in_millions = population / 10^6)
```

We can write `population` rather than `murders$population`. The function `mutate` knows we are grabbing columns from `murders`.

Use the function `mutate` to add a murders column named `rate` with the per 100,000 murder rate as in the example code above. Make sure you redefine `murders` as done in the example code above ( murders <- [your code]) so we can keep using this variable.

```{r}
murders <- mutate(murders, rate = total / population * 100000)
murders
```


2. If `rank(x)` gives you the ranks of `x` from lowest to highest, `rank(-x)` gives you the ranks from highest to lowest. Use the function `mutate` to add a column `rank` containing the rank, from highest to lowest murder rate. Make sure you redefine `murders` so we can keep using this variable.

```{r}
murders <- mutate(murders, rank = rank(-rate))
murders
```


3. With **dplyr**, we can use `select` to show only certain columns. For example, with this code we would only show the states and population sizes:

```{r}
select(murders, state, population) %>% head()
```

Use `select` to show the state names and abbreviations in `murders`. Do not redefine `murders`, just show the results.

```{r}
select(murders , state, abb)
```


4. The **dplyr** function `filter` is used to choose specific rows of the data frame to keep. Unlike `select` which is for columns, `filter` is for rows. For example, you can show just the New York row like this:

```{r}
filter(murders, state == "New York")
```

You can use other logical vectors to filter rows.

Use `filter` to show the top 5 states with the highest murder rates. After we add murder rate and rank, do not change the murders dataset, just show the result. Remember that you can filter based on the `rank` column.

```{r}
filter(murders, rank <= 5)
```


5. We can remove rows using the `!=` operator. For example, to remove Florida, we would do this:

```{r}
no_florida <- filter(murders, state != "Florida")
```

Create a new data frame called `no_south` that removes states from the South region. How many states are in this category? You can use the function `nrow` for this.

```{r}
no_south <- filter(murders, region != "South")
nrow(no_south)
```

**34**

6. We can also use `%in%` to filter with **dplyr**. You can therefore see the data from New York and Texas like this:

```{r}
filter(murders, state %in% c("New York", "Texas"))
```

Create a new data frame called `murders_nw` with only the states from the Northeast and the West. How many states are in this category?

```{r}
murders_nw <- filter(murders, region %in% c("Northeast", "West"))
nrow(murders_nw)
```

**22**


7. Suppose you want to live in the Northeast or West **and** want the murder rate to be less than 1. We want to see the data for the states satisfying these options. Note that you can use logical operators with `filter`. Here is an example in which we filter to keep only small states in the Northeast region.

```{r}
filter(murders, population < 5000000 & region == "Northeast")
```

Make sure `murders` has been defined with `rate` and `rank` and still has all states. Create a table called `my_states` that contains rows for states satisfying both the conditions: it is in the Northeast or West and the murder rate is less than 1. Use `select` to show only the state name, the rate and the rank.

```{r}
my_states <- filter(murders, (region == "Northeast" | region == "West") & rate < 1)
select(my_states, state, rate, rank)
```



## 5.5 The pipe: `%>%`

With *dplyr* we can perform a series of operations, for example `select` and then `filter`, by sending the results of one function to another using what is called the pipe operator: `%>%`. Some details are included below.

In *dplyr* we can write code that looks more like a description of what we want to do without intermediate objects:

original data → select → filter

For such an operation, we can use the pipe `%>%`. The code looks like this:

```{r}
murders %>% select(state, region, rate) %>% filter(rate <= 0.71)
#>           state        region  rate
#> 1        Hawaii          West 0.515
#> 2          Iowa North Central 0.689
#> 3 New Hampshire     Northeast 0.380
#> 4  North Dakota North Central 0.595
#> 5       Vermont     Northeast 0.320
```

In general, the pipe sends the result of the left side of the pipe to be the first argument of the function on the right side of the pipe. Here is a very simple example:

```{r}
16 %>% sqrt()
#> [1] 4
```

We can continue to pipe values along:

```{r}
16 %>% sqrt() %>% log2()
#> [1] 2
```

The above statement is equivalent to `log2(sqrt(16))`.

Remember that the pipe sends values to the first argument, so we can define other arguments as if the first argument is already defined:

```{r}
16 %>% sqrt() %>% log(base = 2)
#> [1] 2
```

Therefore, when using the pipe with data frames and **dplyr**, we no longer need to specify the required first argument since the *dplyr** functions we have described all take the data as the first argument. In the code we wrote:

```{r}
murders %>% select(state, region, rate) %>% filter(rate <= 0.71)
```

`murders` is the first argument of the `select` function, and the new data frame (formerly `new_table`) is the first argument of the `filter` function.

Note that the pipe works well with functions where the first argument is the input data. Functions in *tidyverse** packages like *dplyr** have this format and can be used easily with the pipe.



## 5.6 Exercises

1. The pipe `%>%` can be used to perform operations sequentially without having to define intermediate objects. Start by redefining murder to include rate and rank.

```{r}
murders <- mutate(murders, rate =  total / population * 100000, rank = rank(-rate))
```

In the solution to the previous exercise, we did the following:

```{r}
my_states <- filter(murders, region %in% c("Northeast", "West") & rate < 1)

select(my_states, state, rate, rank)
```

The pipe `%>%` permits us to perform both operations sequentially without having to define an intermediate variable `my_states`. We therefore could have mutated and selected in the same line like this:

```{r}
mutate(murders, rate =  total / population * 100000, rank = rank(-rate)) %>%
  select(state, rate, rank)
```

Notice that `select` no longer has a data frame as the first argument. The first argument is assumed to be the result of the operation conducted right before the `%>%`.

Repeat the previous exercise, but now instead of creating a new object, show the result and only include the state, rate, and rank columns. Use a pipe `%>%` to do this in just one line.

```{r}
filter(murders, (region == "Northeast" | region == "West") & rate < 1) %>% select(state, rate, rank)
```


2. Reset `murders` to the original table by using `data(murders)`. Use a pipe to create a new data frame called `my_states` that considers only states in the Northeast or West which have a murder rate lower than 1, and contains only the state, rate and rank columns. The pipe should also have four components separated by three `%>%`. The code should look something like this:

```{r}
my_states <- murders %>%
  mutate SOMETHING %>% 
  filter SOMETHING %>% 
  select SOMETHING
```

```{r}
data(murders)
my_states <- murders %>%
  mutate(rate = total / population * 100000, rank = rank(-rate)) %>%
  filter((region == "Northeast" | region == "West") & rate < 1) %>%
  select(state, rate, rank)
```


## 5.7 Summarizing data

An important part of exploratory data analysis is summarizing data. The average and standard deviation are two examples of widely used summary statistics. More informative summaries can often be achieved by first splitting data into groups. In this section, we cover two new *dplyr** verbs that make these computations easier: `summarize` and `group_by`. We learn to access resulting values using the pull function.



### 5.7.1 `summarize`

The `summarize` function in **dplyr** provides a way to compute summary statistics with intuitive and readable code. We start with a simple example based on heights. The `heights` dataset includes heights and sex reported by students in an in-class survey.

```{r}
library(dplyr)
library(dslabs)
data(heights)
```

The following code computes the average and standard deviation for females:

```{r}
s <- heights %>% 
  filter(sex == "Female") %>%
  summarize(average = mean(height), standard_deviation = sd(height))
s
#>   average standard_deviation
#> 1    64.9               3.76
```

This takes our original data table as input, filters it to keep only females, and then produces a new summarized table with just the average and the standard deviation of heights. We get to choose the names of the columns of the resulting table. For example, above we decided to use `average` and `standard_deviation`, but we could have used other names just the same.

Because the resulting table stored in `s` is a data frame, we can access the components with the accessor `$`:

```{r}
s$average
#> [1] 64.9
s$standard_deviation
#> [1] 3.76
```

As with most other dplyr functions, `summarize` is aware of the variable names and we can use them directly. So when inside the call to the `summarize` function we write `mean(height)`, the function is accessing the column with the name “height” and then computing the average of the resulting numeric vector.

We can compute any other summary that operates on vectors and returns a single value. For example, we can add the median, min and max heights like this:

```{r}
heights %>% 
  filter(sex == "Female") %>%
  summarize(median = median(height), minimum = min(height), maximum = max(height))
#>   median minimum maximum
#> 1     65      51      79
```

We can obtain these three values with just one line using the `quantile` function: for example, `quantile(x, c(0,0.5,1))` returns the min (0th percentile), median (50th percentile), and max (100th percentile) of the vector `x`. However, if we attempt to use a function like this that returns two or more values inside `summarize`:

```{r}
heights %>% 
  filter(sex == "Female") %>%
  summarize(range = quantile(height, c(0, 0.5, 1)))
```

we will receive an error: `Error: expecting result of length one, got : 2`. With the function `summarize`, we can only call functions that return a single value.

```{r}
heights %>% 
  filter(sex == "Female") %>%
  summarize(range = quantile(height, c(0)), less = quantile(height, c(1)))
```

하나씩은 됨.

For another example of how we can use the `summarize` function, let’s compute the average murder rate for the United States. Remember our data table includes total murders and population size for each state and we have already used **dplyr** to add a murder rate column:

```{r}
data(murders)
murders <- murders %>% mutate(rate = total/population*100000)
```

Remember that the US murder rate is *not* the average of the state murder rates:

```{r}
summarize(murders, mean(rate))
#>   mean(rate)
#> 1       2.78
```

This is because in the computation above the small states are given the same weight as the large ones. The US murder rate is the total number of murders in the US divided by the total US population. So the correct computation is:

```{r}
us_murder_rate <- murders %>%
  summarize(rate = sum(total) / sum(population) * 100000)
us_murder_rate
#>   rate
#> 1 3.03
```

This computation counts larger states proportionally to their size which results in a larger value.



### 5.7.2 `pull`

The `us_murder_rate` object defined above represents just one number. Yet we are storing it in a data frame:

```{r}
class(us_murder_rate)
#> [1] "data.frame"
```

since, as most *dplyr* functions, `summarize` always returns a data frame.

This might be problematic if we want to use this result with functions that require a numeric value. Here we show a useful trick for accessing values stored in data when using pipes: when a data object is piped that object and its columns can be accessed using the `pull` function. To understand what we mean take a look at this line of code:

```{r}
us_murder_rate %>% pull(rate)
#> [1] 3.03
```

This returns the value in the `rate` column of `us_murder_rate` making it equivalent to `us_murder_rate$rate`.

To get a number from the original data table with one line of code we can type:

```{r}
us_murder_rate <- murders %>% 
  summarize(rate = sum(total) / sum(population) * 100000) %>%
  pull(rate)

us_murder_rate
#> [1] 3.03
```
숫자로 저장함.

which is now a numeric:

```{r}
class(us_murder_rate)
#> [1] "numeric"
```



### 5.7.3 Group then summarize with `group_by`

A common operation in data exploration is to first split data into groups and then compute summaries for each group. For example, we may want to compute the average and standard deviation for men’s and women’s heights separately. The `group_by` function helps us do this.

If we type this:

```{r}
heights %>% group_by(sex)
```

The result does not look very different from `heights`, except we see `Groups: sex [2]` when we print the object. Although not immediately obvious from its appearance, this is now a special data frame called a grouped data frame and *dplyr* functions, in particular `summarize`, will behave differently when acting on this object.

Conceptually, you can think of this table as many tables, with the same columns but not necessarily the same number of rows, stacked together in one object. When we summarize the data after grouping, this is what happens:

```{r}
heights %>% 
  group_by(sex) %>%
  summarize(average = mean(height), standard_deviation = sd(height))
#> # A tibble: 2 x 3
#>   sex    average standard_deviation
#>   <fct>    <dbl>              <dbl>
#> 1 Female    64.9               3.76
#> 2 Male      69.3               3.61
```

The `summarize` function applies the summarization to each group separately.

For another example, let’s compute the median murder rate in the four regions of the country:

```{r}
murders %>% 
  group_by(region) %>%
  summarize(median_rate = median(rate))
#> # A tibble: 4 x 2
#>   region        median_rate
#>   <fct>               <dbl>
#> 1 Northeast            1.80
#> 2 South                3.40
#> 3 North Central        1.97
#> 4 West                 1.29
```



## 5.8 Sorting data frames

When examining a dataset, it is often convenient to sort the table by the different columns. We know about the `order` and `sort` function, but for ordering entire tables, the *dplyr* function `arrange` is useful. For example, here we order the states by population size:

```{r}
murders %>%
  arrange(population) %>%
  head()
#>                  state abb        region population total   rate
#> 1              Wyoming  WY          West     563626     5  0.887
#> 2 District of Columbia  DC         South     601723    99 16.453
#> 3              Vermont  VT     Northeast     625741     2  0.320
#> 4         North Dakota  ND North Central     672591     4  0.595
#> 5               Alaska  AK          West     710231    19  2.675
#> 6         South Dakota  SD North Central     814180     8  0.983
```

With `arrange` we get to decide which column to sort by. To see the states by population, from smallest to largest, we arrange by `rate` instead:
((?? population이라고?))

```{r}
murders %>% 
  arrange(rate) %>% 
  head()
#>           state abb        region population total  rate
#> 1       Vermont  VT     Northeast     625741     2 0.320
#> 2 New Hampshire  NH     Northeast    1316470     5 0.380
#> 3        Hawaii  HI          West    1360301     7 0.515
#> 4  North Dakota  ND North Central     672591     4 0.595
#> 5          Iowa  IA North Central    3046355    21 0.689
#> 6         Idaho  ID          West    1567582    12 0.766
```

Note that the default behavior is to order in ascending order.

In *dplyr*, the function `desc` transforms a vector so that it is in descending order. To sort the table in descending order, we can type:

```{r}
murders %>% 
  arrange(desc(rate)) %>% 
  head()
#>                  state abb        region population total  rate
#> 1 District of Columbia  DC         South     601723    99 16.45
#> 2            Louisiana  LA         South    4533372   351  7.74
#> 3             Missouri  MO North Central    5988927   321  5.36
#> 4             Maryland  MD         South    5773552   293  5.07
#> 5       South Carolina  SC         South    4625364   207  4.48
#> 6             Delaware  DE         South     897934    38  4.23
```



### 5.8.1 Nested sorting

If we are ordering by a column with ties, we can use a second column to break the tie. Similarly, a third column can be used to break ties between first and second and so on. Here we order by region, then within region we order by murder rate:

```{r}
murders %>% 
  arrange(region, rate) %>% 
  head()
#>           state abb    region population total  rate
#> 1       Vermont  VT Northeast     625741     2 0.320
#> 2 New Hampshire  NH Northeast    1316470     5 0.380
#> 3         Maine  ME Northeast    1328361    11 0.828
#> 4  Rhode Island  RI Northeast    1052567    16 1.520
#> 5 Massachusetts  MA Northeast    6547629   118 1.802
#> 6      New York  NY Northeast   19378102   517 2.668
```



### 5.8.2 The top n

In the code above, we have used the function `head` to avoid having the page fill up with the entire dataset. If we want to see a larger proportion, we can use the `top_n` function. This function takes a data frame as it’s first argument, the number of rows to show in the second, and the variable to filter by in the third. Here is an example of how to see the top 10 rows:

```{r}
murders %>% top_n(10, rate)
```

Note that rows are not sorted by `rate`, only *filtered*. If want to sort, we need to use `arrange`. Note that if the third argument is left blank, `top_n`, filters by the last column.



## 5.9 Exercises

For these exercises, we will be using the data from the survey collected by the United States National Center for Health Statistics (NCHS). This center has conducted a series of health and nutrition surveys since the 1960’s. Starting in 1999, about 5,000 individuals of all ages have been interviewed every year and they complete the health examination component of the survey. Part of the data is made available via the **NHANES** package. Once you install the **NHANES** package, you can load the data like this:

```{r}
library(NHANES)
data(NHANES)
```

The **NHANES** data has many missing values. Remember that the main summarization function in R will return `NA` if any of the entries of the input vector is an `NA`. Here is an example:

```{r}
library(dslabs)
data(na_example)
mean(na_example)
#> [1] NA
sd(na_example)
#> [1] NA
```

To ignore the `NA`s we can use the `na.rm` argument:

```{r}
mean(na_example, na.rm = TRUE)
#> [1] 2.3
sd(na_example, na.rm = TRUE)
#> [1] 1.22
```

Let’s now explore the NHANES data.

1. We will provide some basic facts about blood pressure. First let’s select a group to set the standard. We will use 20-29 year old females. `AgeDecade` is a categorical variable with these ages. Note that the category is coded like " 20-29“, with a space in front! What is the average and standard deviation of systolic blood pressure as saved in the `BPSysAve` variable? Save it to a variable called `ref`.

Hint: Use `filter` and `summarize` and use the `na.rm = TRUE` argument when computing the average and standard deviation. You can also filter the NA values using `filter`.



```{r}
ref <- NHANES %>%
  filter(Gender == "female") %>%
  filter(AgeDecade == " 20-29") %>%
  summarize(mean = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
ref
```


2. Using a pipe, assign the average to a numeric variable `ref_avg`. Hint: Use the code similar to above and then `pull`.

```{r}
ref_avg <- NHANES %>%
  filter(Gender == "female") %>%
  filter(AgeDecade == " 20-29") %>%
  summarize(mean = mean(BPSysAve, na.rm = TRUE)) %>%
  pull(mean)
ref_avg
```


3. Now report the min and max values for the same group.

```{r}
ref_ver_2 <- NHANES %>%
  filter(Gender == "female") %>%
  filter(AgeDecade == " 20-29") %>%
  summarize(min = min(BPSysAve, na.rm = TRUE), max = max(BPSysAve, na.rm = TRUE))
ref_ver_2
```
이렇게 하는 거 맞나...?


4. Compute the average and standard deviation for females, but for each age group separately rather than a selected decade as in question 1. Note that the age groups are defined by `AgeDecade`. Hint: rather than filtering by age and gender, filter by `Gender` and then use `group_by`.

```{r}
NHANES %>%
  filter(Gender == "female") %>%
  filter(AgeDecade != "NA") %>%
  group_by(AgeDecade) %>%
  summarize(mean = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
```


5. Repeat exercise 4 for males.

```{r}
NHANES %>%
  filter(Gender == "male") %>%
  filter(AgeDecade != "NA") %>%
  group_by(AgeDecade) %>%
  summarize(mean = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
```


6. We can actually combine both summaries for exercises 4 and 5 into one line of code. This is because `group_by` permits us to group by more than one variable. Obtain one big summary table using `group_by(AgeDecade, Gender)`.

```{r}
NHANES %>%
  filter(AgeDecade != "NA") %>%
  group_by(AgeDecade, Gender) %>%
  summarize(mean = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
```


7. For males between the ages of 40-49, compare systolic blood pressure across race as reported in the `Race1` variable. Order the resulting table from lowest to highest average systolic blood pressure.

```{r}
NHANES %>%
  filter(Gender == "male") %>%
  filter(AgeDecade == " 40-49") %>%
  group_by(Race1) %>%
  summarize(mean = mean(BPSysAve, na.rm = TRUE)) %>%
  arrange(mean)
```



## 5.10 Tibbles



### 5.10.1 Tibbles display better



### 5.10.2 Subsets of tibbles are tibbles



### 5.10.3 Tibbles can have complex entries



### 5.10.4 Tibbles can be grouped



### 5.10.5 Create a tibble using `data_frame` instead of `data.frame`



## 5.11 The dot operator



## 5.12 `do`



## 5.13 The purrr package



## 5.14 Tidyverse conditionals



### 5.14.1 `case_when`



### 5.14.2 `between`



## 5.15 Exercises

1. Load the `murders` dataset. Which of the following is true?

A. `murders` is in tidy format and is stored in a tibble.

B. `murders` is in tidy format and is stored in a data frame.

C. `murders` is not in tidy format and is stored in a tibble.

D. `murders` is not in tidy format and is stored in a data frame.

2. Use `as_tibble` to covert the `murders` data table into a tibble and save it in an object called `murders_tibble`.

3. Use the `group_by` function to convert murders into a tibble that is grouped by region.

4. Write tidyverse code that is equivalent to this code: `exp(mean(log(murders$population)))`. Write it using the pipe so that each function is called without arguments. Use the dot operator to access the population. Hint: The code should start with `murders %>%`.

5. Use the `map_df` to create a data frame with three columns named `n`, `s_n`, and `s_n_2`. The first column should contain the numbers 1 through 100. The second and third columns should each contain the sum of 1 through n with n the row number.